---
title: "Sentiment Analysis over Ebola Data"
author: "Ajinkya Sheth"
date: "May 19, 2019"
output:
  html_document:
    df_print: paged
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.



```{r setup, include=FALSE}
# 
library(zoo)
library(dplyr)
library(pander)
library(tidytext)
library(tm)
library(ggplot2)
library(ggthemes)
library(tidyr)
#library(corpus)
library(wordcloud)
library(hash)
library(sf)
library(tmap)

library(raster)
library(spData)

library(sentimentr)
library(ggmap)

library(gganimate)

library(gifski)
library(transformr)

library(mapproj)

GOOGLEAPI <- 'AIzaSyBB4L0iiLI6vTM3Tg-GbfcFo9lNarNRYU8'
SIERRALEONE_SHP <- read_sf('./SLE_adm3.shp')
```


```{r}
install.packages('epiweek')

```

Data Explorations

```{r}
trigger <- read.csv(file="./Trigger_Other.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)

```

```{r}
concerns=trigger$t_q6
questions=trigger$t_q7
risks=trigger$t_q8
byelaws=trigger$t_q9
discussions=trigger$t_q10
capability=trigger$t_q11

```

```{r}
# Function Testing
add <- function(x,y=1) {
  z <- 1
  x+y+z
}
add(2)
(function(x,y=0) {x-y})(1,5)
```



```{r}
concerns_df <- tibble(line = 1:length(concerns),text=concerns )
tidy_corpus <- concerns_df %>% 
  unnest_tokens(word, text)
data('stop_words')
tidy_corpus <- tidy_corpus %>% 
  anti_join(stop_words)


```


```{r}
tidy_corpus %>%
  count(word, sort = TRUE) %>%
  filter(n > 100) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
    geom_col() +
    xlab(NULL) +
    coord_flip() +
  theme_tufte()
```

Removing 'ebola' because we are trying to find sentiment during ebola outbreak and the word 'ebola' may not add any value to our sentiment analysis. (Check with Laura) Also removing NA

```{r}
tidy_corpus=tidy_corpus[tidy_corpus$word != "ebola",]
tidy_corpus=tidy_corpus[!is.na(tidy_corpus$word),]


```

```{r}
corpus_sentiment <- tidy_corpus %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

word_counts <- tidy_corpus %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

```
```{r}
positive_sentiment <- corpus_sentiment %>%
  filter(sentiment>0) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip() +
  theme_tufte()
```
```{r}
word_counts %>%
  filter(sentiment=='positive')
```


```{r}


layout(matrix(c(1, 2), nrow=2), heights=c(1, 6))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Positive Sentiment")

word_counts %>%
  filter(sentiment=='positive') %>%
  with(wordcloud(word, n, max.words = 50))

layout(matrix(c(1, 2), nrow=2), heights=c(1, 6))
par(mar=rep(0, 4))
plot.new()
text(x=0.5, y=0.5, "Negative Sentiment")
word_counts %>%
  filter(sentiment=='negative') %>%
  with(wordcloud(word, n, max.words = 50))
```

```{r}
positive_words <- word_counts %>%
  filter(sentiment=='positive') %>%
  drop()
```

```{r}
library(hash)

word_dict = hash()

word_sentiment_df <- word_counts %>%
  select(word,sentiment)


#by(word_sentiment_df, 1:nrow(word_sentiment_df), function(row) dostuff)

for (row in 1:nrow(word_sentiment_df)) {
  if (word_sentiment_df[[2]][row] == 'positive') {
    word_dict[[word_sentiment_df[[1]][row]]] <- 1  
  }
  else
    word_dict[[word_sentiment_df[[1]][row]]] <- -1
  
  
}


```


```{r}
calculate_sentiment <- function(list_words) {
  sum=0
  for (word in list_words) {
    if(!is.na(word))
    if(!word == "") {
      if(has.key(word,word_dict)) {
        sum <- sum + word_dict[[word]]
      }
    }
  }
  sum
}

calculate_positive_sentiment <- function(list_words) {
  sum=0
  for (word in list_words) {
    if(!is.na(word))
    if(!word == "") {
      if(has.key(word,word_dict)) {
        if (word_dict[[word]] >=0) {
        sum <- sum + word_dict[[word]]
        }
      }
    }
  }
  sum
}

calculate_negative_sentiment <- function(list_words) {
  sum=0
  for (word in list_words) {
    if(!is.na(word))
    if(!word == "") {
      if(has.key(word,word_dict)) {
        if (word_dict[[word]] <=0) {
        sum <- sum + word_dict[[word]]
        }
      }
    }
  }
  sum
}
```


```{r}

for (row in 1:nrow(concerns_df)) {
  qq <- concerns_df$text[row] %>%
        tolower() %>%
        strsplit(" ") %>%
        unlist
  #print(qq[[1]])
  #print(aa)
  concerns_df$sentiment[row] <- calculate_sentiment(qq)
}

for (row in 1:nrow(concerns_df)) {
  qq <- concerns_df$text[row] %>%
        tolower() %>%
        strsplit(" ") %>%
        unlist
  #print(qq[[1]])
  #print(aa)
  concerns_df$positive_sentiment[row] <- calculate_positive_sentiment(qq)
}

for (row in 1:nrow(concerns_df)) {
  qq <- concerns_df$text[row] %>%
        tolower() %>%
        strsplit(" ") %>%
        unlist
  #print(qq[[1]])
  #print(aa)
  concerns_df$negative_sentiment[row] <- calculate_negative_sentiment(qq)
}


#concerns_df$text %>%
#  tolower() %>%
#  strsplit(" ") %>%
#  apply(1,calculate_negative_sentiment)

concerns_df$Trig_date <- trigger$Trig_date
concerns_df$Trig_date <- as.Date(concerns_df$Trig_date, "%m/%d/%Y")

require(ggplot2)
ggplot( data = concerns_df, aes( Trig_date, sentiment )) + geom_line()
concerns_df <- 1
```

```{r}
trigger
```




sentimentr tests
```{r}
sentences_list <- c("I like ice-cream", 
                    "I do not like ice-cream", 
                    "I hate ice-cream", 
                    "I hardly like ice-cream",
                    "Do you think I like ice-cream?")

sentences_list %>%
  get_sentences() %>%
  sentiment


```

```{r}
concerns_by_day
```

Notes for Snetiment by time:
- Two fields: mean sentiment as a function of a day
- Approx. Positive slope
- Two extremely negative peaks, indicate sentiment about concerns over entire country fell
- Terrible news can be a confounding factor



```{r}
A <-  (trigger %>%
  dplyr::select(Trig_date,District,Chiefdom,t_q6))

B <-  (trigger$t_q6 %>%
         get_sentences() %>%
  sentiment_by(by=NULL))

sentiment_concerns <- cbind(A,B)

sentiment_concerns$Trig_date <- as.Date(sentiment_concerns$Trig_date, "%m/%d/%Y")

concerns_by_day <- sentiment_concerns %>%
  dplyr::select(Trig_date,ave_sentiment)

concerns_by_day <- concerns_by_day %>%
  group_by(Trig_date) %>%
  summarise(mean_ave = mean(ave_sentiment, na.rm = TRUE))

require(ggplot2)
ggplot( data = concerns_by_day, aes( Trig_date, mean_ave )) + geom_line() + ggtitle('Mean Sentiment')

geom_point()

```


```{r}
# tm_shape(65) + tm_fill()
# 
# sentiment_concerns
# 
# world[world$name_long == 'Sierra Leone',]
# 
# register_google(key = GOOGLEAPI)

concerns_by_chiefdom <- sentiment_concerns %>%
  dplyr::select(District, Chiefdom, ave_sentiment)

concerns_by_chiefdom <- concerns_by_chiefdom %>%
  group_by(District,Chiefdom) %>%
  summarise(mean_ave = mean(ave_sentiment, na.rm = TRUE))


concerns_by_district <- sentiment_concerns %>%
  dplyr::select(District, ave_sentiment)

concerns_by_district <- concerns_by_district %>%
  group_by(District) %>%
  summarise(mean_sent = mean(ave_sentiment, na.rm = TRUE))

counts_by_district <- sentiment_concerns %>%
  dplyr::select(District) %>%
  group_by(District) %>%
  summarise(count_sent = n())




write.csv(concerns_by_chiefdom, file = "concerns_by_chiefdom.csv")

SIERRALEONE_SHP

concerns_by_district
```
```{r}


plt <- ggplot()


counts_plot <- district_df %>%
  left_join(counts_by_district,by=c("shp.NAME_2"="District"))
#write.csv(data.frame(shp$admin3Name,shp$admin3RefN,shp$admin2Name),file = "ShapeFile3.csv")

plt + geom_sf(data = counts_plot, aes(fill=count_sent)) + scale_fill_gradient2(low='red',high='green',mid = 'yellow', midpoint = 800)

```


Notes about sentiment by space:
- Freetown and another district missing from the data
- south-western part more negative sentiment
- Measures against ebola may be better in other parts of the country 



```{r}
shp <- read_sf('./shp1/SLE_adm2.shp')

plt <- ggplot()

district_df<-data.frame(shp$NAME_2,shp$geometry)
concerns_plot<-district_df %>%
  left_join(concerns_by_district,by=c("shp.NAME_2"="District"))
#write.csv(data.frame(shp$admin3Name,shp$admin3RefN,shp$admin2Name),file = "ShapeFile3.csv")

plt + geom_sf(data = concerns_plot, aes(fill=mean_sent)) + scale_fill_gradient2(low='red',high='green', mid='yellow',midpoint=-0.075)

#concerns_by_district
#concerns_plot
```
```{r}
concerns_by_time
concerns_plot
```
```{r}
concerns_time_geo <- sentiment_concerns %>%
  dplyr::select(Trig_date,District,ave_sentiment) %>%
  group_by(Trig_date,District) %>%
  summarise(mean_sent = mean(ave_sentiment, na.rm = TRUE))

concerns_time_geo<-district_df %>%
  left_join(concerns_time_geo,by=c("shp.NAME_2"="District"))

concerns_time_geo <- concerns_time_geo[!is.na(concerns_time_geo$Trig_date),]


concerns_time_geo_animate <- ggplot() + geom_sf(data = concerns_time_geo, aes(fill=mean_sent,frame=Trig_date)) + transition_time(Trig_date) + coord_sf() + theme_minimal() + labs(title = "Date: {frame_time}")
## animate plot with gganimate
animate(concerns_time_geo_animate, fps=4) + ease_aes()

concerns_time_geo_animate
  


length(concerns_time_geo$Trig_date)
```

```{r}
Trig_month <- concerns_time_geo$Trig_date %>%
  strftime(format="%m")

concerns_month_geo <- data.frame(concerns_time_geo, Trig_month)

MIN_DATE <- min(concerns_time_geo$Trig_date)

concerns_month_geo$
concerns_month_geo$Trig_month + 1

Trig_month<-(as.yearmon(concerns_time_geo$Trig_date)-as.yearmon(MIN_DATE))*12
concerns_month_geo <- data.frame(concerns_time_geo, Trig_month)

Trig_month

concerns_month_geo

concerns_month_geo <- concerns_month_geo %>%
  dplyr::select(Trig_month,shp.NAME_2,mean_sent) %>%
  group_by(Trig_month,shp.NAME_2) %>%
  summarise(mean_sent = mean(mean_sent, na.rm = TRUE))

concerns_month_geo<-district_df %>%
  left_join(concerns_month_geo,by=c("shp.NAME_2"="shp.NAME_2"))

concerns_month_geo <- concerns_month_geo[!is.na(concerns_month_geo$Trig_month),]


concerns_month_geo_animate <- ggplot() + geom_sf(data = concerns_month_geo, aes(fill=mean_sent,frame=Trig_month)) + transition_time(Trig_month) + coord_sf() + theme_minimal() + labs(title = "Date: {frame_time}")
## animate plot with gganimate
animate(concerns_month_geo_animate, fps=1)


concerns_month_geo$Trig_month

concerns_month_geo

```

```{r}

concerns_dis_mon <- sentiment_concerns %>%
  dplyr::select(Trig_date,District, ave_sentiment)

concerns_dis_mon$Trig_date <- as.Date(as.yearmon(concerns_dis_mon$Trig_date))
#concerns_dis_mon$Trig_date <- as.Date(concerns_dis_mon$Trig_date, "%b %Y")

concerns_dis_mon <- concerns_dis_mon %>%
  group_by(District,Trig_date) %>%
  summarise(mean_sent = mean(ave_sentiment, na.rm = TRUE))

concerns_dis_mon
ggplot( data = concerns_dis_mon, aes(x=Trig_date, y=mean_sent)) + geom_line(aes(color = District), size=2) + ggtitle('Mean Sentiment') + scale_x_date(name="Trig_date", date_labels = "%b %Y", breaks="month" )
```



```{r}

concerns_month_geo <-district_df %>%
  left_join(concerns_dis_mon,by=c("shp.NAME_2"="District"))

concerns_month_geo <- concerns_month_geo[!is.na(concerns_month_geo$Trig_date),]

concerns_month_geo


concerns_month_geo_animate <- ggplot() + geom_sf(data = concerns_month_geo, aes(fill=mean_sent,frame=Trig_date)) + scale_fill_gradient2(low='red',high='green', mid='white',midpoint=0) + transition_states(Trig_date, wrap=TRUE) + coord_sf() + theme_tufte() + labs(title = "Date") 
## animate plot with gganimate
animate(concerns_month_geo_animate, fps=3)

anim_save("animation.gif",animation=last_animation())

concerns_month_geo$Trig_month

concerns_month_geo

```
```{r}
library(rgdal)

shp4 <- read_sf('./shp4/sle_admbnda_adm2_1m_gov_ocha.shp')


SL <- readOGR(dsn = "./shp4/sle_admbnda_adm2_1m_gov_ocha.shp",
                    layer="sle_admbnda_adm2_1m_gov_ocha")
proj4string(SL) <- CRS("+proj=longlat")

SL.df <- merge(fortify(SL), as.data.frame(SL), by.x="id", by.y=0)

SL.df$admin2Name %>%
  unique

concerns_dis_mon
ebola_data <- SL.df %>%
  left_join(concerns_dis_mon,by=c("admin2Name"="District"))
ebola_data
ebola_data_sub <- ebola_data[ebola_data$Trig_date=="2015-01-01",]

## create ggplot object ##
ebola_plot_sub <- ggplot(data = ebola_data_sub, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = mean_sent),color="black") +
  scale_fill_gradient(low = "#fcfbfd", high = "#2d004b") +
  coord_map() + theme_minimal()

## Plot ggplot object

ebola_plot_sub
```
```{r}
ebola_plot_animate <- ggplot() + geom_polygon(data = ebola_data, aes(x = long, y = lat, group=group, frame=Trig_date)) +
  geom_polygon(aes(fill = mean_sent),color="black") +
  scale_fill_gradient(low = "#fcfbfd", high = "#2d004b") +
  coord_map() +
  theme_minimal()

## animate plot with gganimate
animate(ebola_plot_animate, fps=3)
```

