---
title: "Alcohol Consumption"
author: "Ajinkya Sheth"
date: "March 9, 2019"
output: 
  html_document:
    smart: no
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

```{r}
alcohol = read.csv("./Alcohol Use_1516.csv", header=TRUE)
demographic = read.csv("./Demographic_1516.csv", header=TRUE)
depression = read.csv("./Depression_1516.csv", header=TRUE)
drug = read.csv("./Drug Use_1516.csv", header=TRUE)

```
## Variables considered for this analysis - Alcohol table
ALQ120Q - How often drink alcohol over past 12 mos
ALQ120U - # days drink alcohol per wk, mo, yr
ALQ130 - Avg # alcoholic drinks/day - past 12 mos
ALQ141Q - # days have 4/5 drinks - past 12 mos
ALQ141U - # days per week, month, year?
ALQ151 - Ever have 4/5 or more drinks every day?
ALQ160 - # days have 4/5 or more drinks in 2 hrs
***

```{r}

alcohol_clean <- alcohol %>%
  select(SEQN,ALQ120Q,ALQ120U, ALQ130, ALQ141Q, ALQ141U, ALQ151, ALQ160) %>%
  na.omit(cols=seq_along(ALQ120Q, ALQ120U, ALQ130, ALQ141Q, ALQ141U, ALQ151, ALQ160)) %>%
  as.data.frame()

#write.table(alcohol_clean, "./alchohol_clean.csv", sep=",")
  
```


Demographic:
RIAGENDR - Gender
RIDAGEYR - Age in years at screening
DMQMILIZ - Served active duty in US Armed Forces
DMDCITZN - Citizenship status
DMDEDUC3 - Education level - Children/Youth 6-19
INDFMIN2 - Annual family income
INDFMPIR - Ratio of family income to poverty
DMDEDUC2 - Education level - Adults 20+
DMDMARTL - Marital status
***

```{r}

demographc_clean <- demographic %>%
  select(SEQN, RIAGENDR, RIDAGEYR, DMQMILIZ, DMDCITZN, DMDEDUC3, INDFMIN2, INDFMPIR, DMDEDUC2, DMDMARTL) %>%
  as.data.frame()
```

Depression:
DPQ010 - Have little interest in doing things
DPQ020 - Feeling down, depressed, or hopeless
DPQ030 - Trouble sleeping or sleeping too much
DPQ040 - Feeling tired or having little energy
DPQ050 - Poor appetite or overeating
DPQ060 - Feeling bad about yourself
DPQ070 - Trouble concentrating on things
DPQ080 - Moving or speaking slowly or too fast
DPQ090 - Thought you would be better off dead
DPQ100 - Difficulty these problems have caused
***

## Quantify depression 
We quantify depression as a sum of all the variables in the depression table except DPQ100 
Since DPQ100 by nature and by definition is a multiplicative variable


```{r}
depression_clean <- depression %>%
  na.omit(seq_along(DPQ010, DPQ020, DPQ030, DPQ040, DPQ050, DPQ060, DPQ070, DPQ080, DPQ090, DPQ100)) %>%
  as.data.frame()

cols_to_mutate <- c("DPQ010", "DPQ020", "DPQ030", "DPQ040", "DPQ050", "DPQ060", "DPQ070", "DPQ080", "DPQ090")

depression_clean %>% 
  select(DPQ010:DPQ100) %>% 
    mutate_at(cols_to_mutate, function(x) {
    case_when(
      x == 1 ~ 1,
      x == 2 ~ 2,
      x == 3 ~ 3,
      x == 0 ~ 0,
      T ~ as.numeric(NA)
    )
  }) %>%
  rowSums(na.rm=TRUE) -> depression_clean$DepressionScore 


# We scraped out the multiplicative variable in our final analysis
normalizeDepressionScore <- function(score, factor) {
  case_when(
      factor == 0 ~ 1*score,
      factor == 1 ~ 2*score,
      factor == 2 ~ 4*score,
      factor == 3 ~ 8*score,
      T ~ as.numeric(NA)
  )
}

#\depression_clean$DepressionScore <- normalizeDepressionScore(depression_clean$DepressionScore,                                                             depression_clean$DPQ100)
  
```

## Join alcohol, demographic and drepression table for univariate analysis
According to CDC definitions, excess consumption of alcohol is defined differently for males and females.
Hence, we need gender information to infer relationship between alcoholism and depression precisely.
To keep things simple, we are joining demographic and depression table to alcohol table via inner join.
The reason of choosing inner join is because:
We would need all the information to conduct further analysis

```{r}
depression_trunc <- depression_clean %>% select(SEQN,DepressionScore,DPQ090) %>% as.data.frame()

AlcoholAnalysis <- alcohol_clean %>% 
                    inner_join(depression_trunc,
                      by="SEQN",
                      copy=False) %>%
                    inner_join(demographc_clean,
                      by="SEQN",
                      copy=False)
```


## Quantify alcoholism
### Alcohol in the USA
In the United States, a standard drink contains 0.6 ounces (14.0 grams or 1.2 tablespoons) of pure alcohol. Generally, this amount of pure alcohol is found in
12-ounces of beer (5% alcohol content).
8-ounces of malt liquor (7% alcohol content).
5-ounces of wine (12% alcohol content).
1.5-ounces of 80-proof (40% alcohol content) distilled spirits or liquor (e.g., gin, rum, vodka, whiskey).4
(https://www.cdc.gov/alcohol/fact-sheets/alcohol-use.htm)

### 1. Drinkers
**Definition of Moderate Drinking:**
The Dietary Guidelines for Americans defines moderate drinking as up to 1 drink per day for women and up to 2 drinks per day for men. In addition, the Dietary Guidelines do not recommend that individuals who do not drink alcohol start drinking for any reason.
(https://www.cdc.gov/alcohol/fact-sheets/alcohol-use.htm)
**Scheme of quantification:**
Total Alcohol Consumption (TotalConsumption) = Drinks/Day * Drinking days
TotalConsumption = Drinking Days[function(ALQ120Q, ALQ120U)] * Avg drinks/day [ALQ130]


### 2. Binge Drinkers 
**Definition of Binge Drinking:**
The National Institute on Alcohol Abuse and AlcoholismExternal defines binge drinking as a pattern of drinking that brings a person’s blood alcohol concentration (BAC) to 0.08 grams percent or above. This typically happens when men consume 5 or more drinks or women consume 4 or more drinks in about 2 hours.
(https://www.cdc.gov/alcohol/fact-sheets/binge-drinking.htm)
In our analysis, we will consider consumption in throughout the day as well, 
**Scheme of quantification:**
Degree of binge drinking= Number of bringe drinking sessions*Alcohol consumed in bringe drinking sessions


### 3. Heavy Drinkers:
**Definition of Heavy Drinking:**
Heavy drinking is defined as consuming
For women, 8 or more drinks per week.
For men, 15 or more drinks per week.
(https://www.cdc.gov/alcohol/fact-sheets/alcohol-use.htm)
**Scheme of quantification:**
Degree of heavy drinking= Number of heavy drinking sessions * alcohol consumed in heavy drinking sessions

### 4. Recent Excess Consumption:
We consider the field "ALQ160" to establish this metric. 

*Note: Occasional drinkers or drinkers who do not fit into the above category are not conisdered in this analysis*
*Heavy drinkers are out of scope as well*
  
```{r}

getTotalConsumption <- function(freq, unit, avg_drinksByDays){
      case_when(
          freq<375 | unit == 1 ~ freq*7*12*avg_drinksByDays,
          freq<375 | unit == 2 ~ freq*12*avg_drinksByDays,
          freq<375 | unit == 3 ~ freq*avg_drinksByDays,
          freq<375 | unit == 7 ~ as.numeric(NA),
          freq<375 | unit == 9 ~ as.numeric(NA),
          TRUE ~ as.numeric(NA)
          )
 }
 
AlcoholAnalysis$TotalConsumption <- mapply(getTotalConsumption,
                                          as.numeric(AlcoholAnalysis$ALQ120Q), 
                                          as.numeric(AlcoholAnalysis$ALQ120U), 
                                          as.numeric(AlcoholAnalysis$ALQ130))

bingeNumber <- function(gender) {
  case_when (
    gender == 1 ~ 5,
    gender == 2 ~ 4,
    TRUE ~ as.numeric(NA)
  )
}
getBingeConsumption <- function(freq, unit, bingenum){
      case_when(
          freq<375 | unit == 1 ~ freq*7*12*bingenum,
          freq<375 | unit == 2 ~ freq*12*bingenum,
          freq<375 | unit == 3 ~ freq*bingenum,
          freq<375 | unit == 7 ~ as.numeric(NA),
          freq<375 | unit == 9 ~ as.numeric(NA),
          TRUE ~ as.numeric(NA)
          )
}
AlcoholAnalysis$BingeConsumption <- mapply(getBingeConsumption,
                                           AlcoholAnalysis$ALQ141Q,
                                           AlcoholAnalysis$ALQ141U,
                                           bingeNumber(AlcoholAnalysis$RIAGENDR))

getRecentAddiction <- function(freq, bingenum) {
  case_when(
    freq <= 19 ~ freq*bingenum,
    freq == 20 ~ 20*bingenum,
    TRUE ~ as.numeric(NA)
  )
}

AlcoholAnalysis$RecentAddiction <- mapply(getRecentAddiction,
                                           AlcoholAnalysis$ALQ160,
                                           bingeNumber(AlcoholAnalysis$RIAGENDR))


```


### Drop unncessary alcohol columns

```{r}

AlcoholAnalysis <- within(AlcoholAnalysis, rm(ALQ120Q,ALQ120U,ALQ130,ALQ141Q,ALQ141U,ALQ151,ALQ160))
```
```{r}
write.csv(AlcoholAnalysis, file = "AlcoholAnalysis.csv")
```

### Univariate Linear Regression
Depression Score ~ Total Alcohol Consumption
```{r}
mod <- lm(AlcoholAnalysis$DepressionScore ~ AlcoholAnalysis$TotalConsumption, data = AlcoholAnalysis)
#mod <- glm(DPQ090 ~ TotalConsumption,  data = AlcoholAnalysis ) #, family="binomial")
summary(mod)
```
```{r}
plot(mod$model$`AlcoholAnalysis$TotalConsumption`,mod$model$`AlcoholAnalysis$DepressionScore`, xlim=c(0,8000)) # the limit can be set between (5000 and 20000)
abline(mod)

```
```{r}
modBinge <- lm(AlcoholAnalysis$DepressionScore ~ AlcoholAnalysis$BingeConsumption, data = AlcoholAnalysis)
summary(modBinge)
```
```{r}
plot(modBinge$model$`AlcoholAnalysis$BingeConsumption`,modBinge$model$`AlcoholAnalysis$DepressionScore`, xlim=c(0,8000)) # the limit can be set between (5000 and 20000)
abline(modBinge)

```
```{r}
modRecent <- lm(AlcoholAnalysis$DepressionScore ~ AlcoholAnalysis$RecentAddiction, data = AlcoholAnalysis)
summary(modRecent)
```

```{r}
plot(modRecent$model$`AlcoholAnalysis$RecentAddiction`,modRecent$model$`AlcoholAnalysis$DepressionScore`, xlim=c(0,100)) # the limit can be set between (5000 and 20000)
abline(modRecent)

resrd <- resid(modRecent)
hist(resrd)

```
Correlation of Depression score against demographics - age, gender, income
```{r}
#hist(AlcoholAnalysis$DepressionScore)
DepressionDemographics <- demographc_clean %>% 
                    inner_join(depression_trunc,
                      by="SEQN",
                      copy=False)
#head(DepressionDemographics)
modDepDemo <- lm(DepressionScore ~ INDFMIN2, data = DepressionDemographics)
plot(modDepDemo$model$INDFMIN2,modDepDemo$model$DepressionScore, xlim=c(0,20))
abline(modDepDemo)
summary(modDepDemo)
```

```{r}
ServerelyDepressed <- AlcoholAnalysis %>% filter(TotalConsumption >=1500) %>% filter(DepressionScore >= 5)
AlcoholAnalysisFiltered <- AlcoholAnalysis %>% inner_join(ServerelyDepressed)

modFiltered <- lm(AlcoholAnalysisFiltered$DepressionScore ~ AlcoholAnalysisFiltered$TotalConsumption, data = AlcoholAnalysis)
#mod <- glm(DPQ090 ~ TotalConsumption,  data = AlcoholAnalysis ) #, family="binomial")
summary(modFiltered)
```

```{r}
plot(modFiltered$model$`AlcoholAnalysisFiltered$TotalConsumption`,modFiltered$model$`AlcoholAnalysisFiltered$DepressionScore`, ylim=c(0,20),xlim=c(0,10000)) # the limit can be set between (5000 and 20000)
abline(modFiltered)
```


